---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(rulefit)
library(dplyr)
data(titanic, package="binnr2")
```

## Installation

```{r install, eval=FALSE}
devtools::install_git("https://GravesEE@gitlab.ins.risk.regn.net/minneapolis-r-packages/rulefit.git")
```

## Usage

### Creating a RuleFit Model

A RuleFit model uses a tree ensemble to generate its rules. As such, a tree 
ensemble model must be provided to the rulefit function. This funciton returns
a RuleFit object which can be used to mine rules and train rule ensembles.

```{r constructor, warning=FALSE}

mod <- gbm.fit(titanic[-1], 
               titanic$Survived, 
               distribution="bernoulli",
               interaction.depth=3, 
               shrinkage=0.1, 
               verbose = FALSE)

rf <- rulefit(mod, 
              n.trees=100)

print(rf)
```

the `rulefit` function wraps a gbm model in a class that manages rule
construction and model fitting. The rules are generated immediately but the
model is not fit until the `train` function is called.

```{r rules}
head(lapply(rf$rules, toString))
```

For ease of programming *every* internal node is generated -- even the root
node. That is why the first rule listed above is empty. Root nodes are not
splits. This was a design decision and does not affect how the package is used
in practice.

### Training

Training a RuleFit model is as easy as calling the train method. The train
method uses the `cv.glmnet` function from the `glmnet` package and accepts all
of the same arguments.

##### Common Arguments
Argument | Purpose
---|---
x | Dataset of predictors that should match what was used for training the ensemble.
y | Target variable to train against.
family | What is the distribution of the target? Binomial for 0/1 variables.
keep | TRUE/FALSE for whether to keep the out-of-fold predictions. Useful for validation.
alpha | Penatly mixing parameter. LASSO regression uses the default of 0.
nfolds | How many k-folds to train the model with. Defaults to 5.
dfmax | How many variables should the final model have?
parallel | TRUE/FALSE to build kfold models in parallel. Require a backend.

```{r train}
foldids <- sample(1:5, nrow(titanic), replace = T)
rf_ANOVA <- train(rf, titanic[-1], y = titanic$Survived, family="binomial", keep=T, foldid = foldids)
rf_ANCOVA <- train(rf, titanic[-1], y = titanic$Survived, family='binomial', keep=T, linear_components = c('Age', 'Fare'), interact = T, foldid = foldids)
```


### Predicting

Once a RuleFit model is trained. Predictions can be produced by calling the 
predict method. As with the train function, `predict` also takes arguments
accepted by `predict.cv.glmnet`. The most important of which is the lambda
parameter, `s`. The default is to use `s="lambda.min"` which minimizes the
out-of-fold error.

Both a score as well as a sparse matrix of rules can be predicted.

```{r, predict-basics, warning=FALSE}

p_rf_ANOVA <- predict(rf_ANOVA, newx = titanic[-1], s="lambda.1se")
p_rf_ANCOVA <- predict(rf_ANCOVA, titanic[-1], s='lambda.1se')

head(p_rf_ANOVA)
```

The out-of-fold predictions can also be extracted if the model was trained with
`keep=TRUE`. Again, this is working with the `cv.glmnet` API. There is nothing
magical going on here:

#### Comparing RuleFit dev & val to GBM

```{r predict, warning=FALSE}

p_gbm <- predict(mod, titanic[-1], n.trees = gbm.perf(mod, plot.it = F))

roc_rf_ANOVA <- pROC::roc(titanic$Survived, -p_rf_ANOVA)
roc_rf_ANCOVA <- pROC::roc(titanic$Survived, p_rf_ANCOVA)
roc_gbm <- pROC::roc(titanic$Survived, -p_gbm)

plot(roc_rf_ANOVA)
par(new=TRUE)
plot(roc_rf_ANCOVA, col="blue")
par(new=TRUE)
plot(roc_gbm, col="red")

ModelMetrics::auc(actual = titanic$Survived, p_rf_ANOVA)
ModelMetrics::auc(actual = titanic$Survived, p_rf_ANCOVA)
ModelMetrics::auc(actual = titanic$Survived, p_gbm)

```

### Rule Summary

RuleFit also provides a summary method to inspect and measure the coverage of
fitted rules.

```{r, summary}

s <- summary(rf, s="lambda.1se", dedup=TRUE)
DT::datatable(s)

```
